Detecting social information in a dense dataset \\ of infants’ natural visual experience

### 2020-01-30 paper/cogsci/paper_jan28.Rmd
latest updates to cogsci 2020

---
title: "Detecting social information in a dense dataset of infants’ natural visual experience"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

Here, we examine the social information present in the infant visual perspective--the presence of faces and hands--by analyzing a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYcam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008, @smith2015contributions], they may fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

@@

## plot centers of the bounding boxes made for child and adult hands
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ggtitle('A. Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  # theme_few(base_size=12) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  # theme_few(base_size=12) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ggtitle('B. Adult hand density')
```

@@

We suspected that this was in part because children’s own hands were often in view of the camera and unconnected to a pose—-a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s and adult’s hands. Overall, we found that `r round(mean(hands_missed$child_hand_seg),2)` percent of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve to `r hand_performance_adults[3]` somewhat. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in (see Figure \ref{fig:handLocation}).

@@

As children's hands tended to occur in the lower visual field (see Figure 2), we thus re-analyzed the entire dataset while restricting our analysis to the upper 60 percent of the field of view. When we excluded detections that occurred in the bottom 40% of the frame, this decreased the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)`percent to `r round(mean(d_cropped$hand_detected),2)` percent, but only decreased face detections from `r round(mean(d_all$face_openpose),2)` to `r round(mean(d_cropped$face_detected),2)`. Nonetheless, within this modified field of view, we still observed, on average, more hand detections than face detections.

@@

```{r DetByLocation, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3.5, fig.cap = "Proportion of faces and hands by location in which egocentric videos were filmed; data from individual children are plotted separately (location annotations were not available for Y). Data are only taken from the top 60 percent of the frames so as to minimize the contribution of children’s own hands. Each dot represents data from a week in which videos were filmed and are scaled by the number of frames in that datapoint."}

@@

Indeed, a coarse analysis based the location the videos were filmed in already highlights the variability of the the social information in view, both across locations and between individual children. Even within a given, well-defined context---e.g., mealtime in kitchens---there was considerable variability. Indeed, there are at least three ways to feed a young child in a kitchen: 1) sitting in front of the child, facing them as they sit in a high chair; 2) sitting behind the child, holding them as they face outward, and 3) sitting side by side. Each of these positions offer the child differing degrees of visual access to face and hands, and likely tend to vary across individuals

Thus, these analyses underscore how the questions of how, when, from whom, and what data we sample become critically important when we attempt to draw conclusions about the regularities of experience. Further, while unprecedented in size, this dataset still has many limitations. These videos only represent a small portion of the everyday experience of these three children, all of whom come from relatively privileged households in western societies and thus are not representative of the global population. Any idiosyncrasies in how and when these particular families chose to film these videos undoubtedly influences the variability seen here. Finally, without eye-tracking data, we still do not know if children are attending to the social information in their visual field.

Nonetheless, we believe that these advances in methodologies and datasets are a step in the right direction. The present paper demonstrates the feasibility of using a modern, off-the-shelf computer vision model to annotate the entirety of a very large dataset (here, >40M million frames) for the presence and size of people, hands, and faces, representing orders of magnitude more data relative to prior work [e.g., @Fausey2016]. We propose that the large-scale analysis of these kinds of dense datasets—-collected with different fields of view, cameras and from many different laboratories—-have the potential to create generalizable conclusions about the regularities of infant experience that scaffold learning.

>>>>

---
title: "Detecting social information in a dense dataset \\ of infants’ natural visual experience"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

Here, we examine the social information present in the infant visual perspective--the presence of faces and hands--by analyzing a longitudinal collection of over 1400 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYcam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYcam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008; @smith2015contributions], they may fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

@@

## plot centers of the bounding boxes made for child and adult hands
child_hand_plot <- ggplot(child_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ggtitle('A. Child hand density') +
  ylim(0,480) +
  xlim(0,640) +
  ylab('') +
  xlab('') +
  theme_few(base_size=8) +
  theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse()
##
adult_hand_plot <- ggplot(adult_hands, aes(x=center_x, y=center_y)) +
  geom_point(alpha=.1) + 
  stat_density_2d(aes(fill = ..level..), geom="polygon", alpha=.8) +
  coord_fixed(ratio=1) +
  ylim(0,480) +
  xlim(0,640) +
  scale_y_reverse() +
  ylab('') +
  xlab('') +
  theme_few(base_size=8) +
    theme(legend.position="none", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ggtitle('B. Adult hand density')
```

@@

We suspected that this was in part because children’s own hands were often in view of the camera and unconnected to a pose—-a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for a subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s and adult’s hands. Overall, we found that `r round(mean(hands_missed$child_hand_seg),2)*100` percent of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve somewhat to`r round(hand_performance_adults[3],2)`.  We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in (see Figure \ref{fig:handLocation}).

@@

```{r}
## replace NAs for computing prop hands in cropped anlaysis
d_cropped <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) 
```

As children's hands tended to occur in the lower visual field (see Figure 2), we thus re-analyzed the entire dataset while restricting our analysis to the upper 60 percent of the field of view. When we excluded detections that occurred in the bottom 40% of the frame, this decreased the proportion of hand detections from `r round(mean(d_all$hand_openpose),2)*100`% to `r round(mean(d_cropped$hand_detected),2)*100`%, but only decreased face detections from `r round(mean(d_all$face_openpose),2)*100`% to `r round(mean(d_cropped$face_detected),3)*100` % Nonetheless, within this modified field of view, we still observed, on average, more hand detections than face detections (see Figure \ref{fig:FacesAndHands}).

@@

```{r DetByLocation, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of faces and hands by location in which egocentric videos were filmed; data from individual children are plotted separately (location annotations were not available for Y). Data are only taken from the top 60 percent of the frames so as to minimize the contribution of children’s own hands. Each dot represents data from a week in which videos were filmed and are scaled by the number of frames in that datapoint."}

@@

Indeed, a coarse analysis based the location the videos were filmed in already highlights the variability of the the social information in view, both across locations and between individual children. Even within a given, well-defined context---e.g., mealtime in kitchens---there was considerable variability. Indeed, there are at least three ways to feed a young child in a kitchen: 1) sitting in front of the child, facing them as they sit in a high chair; 2) sitting behind the child, holding them as they face outward, and 3) sitting side by side. Each of these positions offer the child differing degrees of visual access to face and hands. Thus, even while the social information in view could be variable across children in different activitiy contexts, they could be stable within children's typical experience.

Thus, these analyses underscore the questions of how, when, from whom, and what data we sample; these choices become critically important when we attempt to draw conclusions about the regularities of experience. Further, while unprecedented in size, this dataset still has many limitations. These videos only represent a small portion of the everyday experience of these three children, all of whom come from relatively privileged households in western societies and thus are not representative of the global population. Any idiosyncrasies in how and when these particular families chose to film these videos undoubtedly influences the variability seen here. Finally, without eye-tracking data, we do not know if children are attending to the social information in their visual field.

Nonetheless, we believe that these advances in methodologies and datasets are a step in the right direction. The present paper demonstrates the feasibility of using a modern, off-the-shelf computer vision model to annotate the entirety of a very large dataset (here, >40M million frames) for the presence and size of people, hands, and faces, representing orders of magnitude more data relative to prior work [e.g., @Fausey2016]. We propose that the large-scale analysis of these kinds of dense datasets, collected with different fields of view, cameras and from many different laboratories, have the potential to create generalizable conclusions about the regularities of infant experience that scaffold learning.
